---
title: "MATH 216 Homework 4"
author: "WRITE YOUR NAME HERE"
output: html_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(maps))
suppressPackageStartupMessages(library(rgdal))
suppressPackageStartupMessages(library(maptools))
```


## Admistrative:

Please indicate

* Who you collaborated with:
* Roughly how much time you spent on this HW:
* What gave you the most trouble:
* Any comments you have: 





## Question 1:

Download the results of the 2000 election from the [School of Public Affairs at 
American University in DC](http://www.american.edu/spa/ccps/Data-Sets.cfm) and 
create a map involving only the lower 48 states that show the proportion of
people who voted for Bush at a **county** level where

* Counties are filled with red when they tend to favor Bush
* Counties are filled with blue when they tend to not favor Bush
* Counties are white with purple when they tend to be split

Which states exhibit the greatest **within state heterogeneity** in voting? 
Come up with a mathematical justification.

Notes:

* Hint: `scale_fill_gradient2(name="", low="blue", high="red", mid="white")` for
the appropriate "mid" point.  See the ggplot2 webpage for this command for
inspiration.
* I believe the county map is from 2010, whereas the election data is from 2000,
as such certain census tracts will not match up. The number is not that large.
* [Counties in Virginia](http://en.wikipedia.org/wiki/List_of_counties_in_Virginia)
* The following code should help

```{r}
# This function eliminates all non-alphanumeric characters and spaces and 
# converts all text to lower case.
clean_text <- function(text){
  text <- gsub("[^[:alnum:]]", "", text)
  text <- gsub(" ", "", text)
  text <- tolower(text)
  return(text)
}
clean_text("HeLLO wOrLd.")

# State and county map of US in 2010
US_state <- map_data("state") %>% 
  tbl_df()
US_county <- map_data("county") %>% 
  tbl_df()
ggplot(data=NULL, aes(x=long, y=lat, group=group)) +
  geom_polygon(data=US_county, fill="white") +
  geom_path(data=US_county, col="black", size=0.01) +
  geom_path(data=US_state, col="black", size=0.05) +
  coord_map()
```




## Question 2:

Make an interactive Shiny app and using

* [Social Explorer](http://www.socialexplorer.com/) to get census data. If you
are stuck with this, please speak to me. The idea is to do this once.
* 

Type `getwd()` in the R console

```{r}
directory_name <- "/Users/aykim/Documents/Teaching/MATH216/HW/HW-4"
shapefile_name <- paste(directory_name, "/VT_census_tracts/tl_2015_50_tract.shp", sep="")
VT <- readOGR(shapefile_name, layer = "tl_2015_50_tract", verbose = FALSE)
```

