---
title: "MATH 216 Homework 4"
author: "WRITE YOUR NAME HERE"
output: html_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For data manipulation and visualization
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(knitr))
# For US county and state maps
suppressPackageStartupMessages(library(maps))
# For loading in shapefiles
suppressPackageStartupMessages(library(rgdal))
suppressPackageStartupMessages(library(maptools))
# For interactive maps
suppressPackageStartupMessages(library(leaflet))
```

## Admistrative:

Please indicate

* Who you collaborated with:
* Roughly how much time you spent on this HW:
* What gave you the most trouble:
* Any comments you have: 





## Question 1:

```{r, cache=TRUE, echo=FALSE}
# Load state and county map of US in 2010 from the maps package and convert them
# to data frames so that we can ggplot them.
US_state <- map_data("state") %>% 
  tbl_df()
US_county <- map_data("county") %>% 
  tbl_df()
```

### Choropleth Map of US Voter Behavior in 2000

Download the results of the 2000 election from the [School of Public Affairs at 
American University in DC](http://www.american.edu/spa/ccps/Data-Sets.cfm) and 
create a map involving only the lower 48 states that show voter behavior at a 
**county** level. To keep things simple let's only consider George W. Bush, Al
Gore, and Ralph Nader. Your write-up must include:

1. A [choropleth map](https://en.wikipedia.org/wiki/Choropleth_map) where
    * Counties are filled with red when they tend to favor Bush
    * Counties are filled with white when they tend to be split
    * Counties are filled with blue when they tend to favor Gore and Nader
2. An answer to the following question: which states exhibit the greatest
**within state heterogeneity** in voting? Come up with a mathematical
justification.


### Hints

* Consider `scale_fill_gradient2(name="", low="blue", high="red", mid="white")` 
for the appropriate "mid" point.  See the ggplot2 webpage for [this
command](http://docs.ggplot2.org/0.9.3.1/scale_gradient2.html) for inspiration.
* I believe the county map is from 2010, whereas the election data is from 2000,
as such certain counties will not match up. The number is not that large.
* [Counties in
Virginia](http://en.wikipedia.org/wiki/List_of_counties_in_Virginia)


### Code Hints

This function eliminates all non-alphanumeric characters and spaces and converts
all text to lower case:

```{r}
clean_text <- function(text){
  text <- gsub("[^[:alnum:]]", "", text)
  text <- gsub(" ", "", text)
  text <- tolower(text)
  return(text)
}
clean_text("HeLLO wOrLd.")
```

For the following `ggplot`, see how I didn't define `data` nor any `aes`thetics
in the base `ggplot`, but only within the respective `geom`'s. This is a nice
trick that allows you to have different `data` and different `aes` apply to
different `geom`'s; in this case state-level vs county-level data.

```{r}
ggplot(data=NULL) +
  geom_polygon(data=US_county, aes(x=long, y=lat, group=group, fill=1)) +
  geom_path(data=US_county, aes(x=long, y=lat, group=group), col="black", size=0.05) +
  geom_path(data=US_state, aes(x=long, y=lat, group=group), col="black", size=0.1) +
  coord_map()
```


### Write-Up

```{r, fig.width=12, fig.height=6}

```






## Question 2:

In this question, you must make an interactive "Single File" Shiny app that uses
Leaflet. For all 184 census tracts in VT in the 2010 census, present
information on the proportion of the population that is either

* White
* African-American
* Hispanic
* Asian and Pacific Island
* Other, including those of two or more races

Use [Social Explorer](http://www.socialexplorer.com/) to get census data. I did
a demo of this in class. If you don't remember how or are stuck, please speak to
me or get help from your peers. Do not submit copies of the same file.

There should be some mechanism in your Shiny app that allows one the user to toggle
between the different ethnic groups.


### Loading Shapefile Data

Here is some starter code:

```{r}
shapefile_name <- paste(getwd(), "/VT_census_tracts/tl_2015_50_tract.shp", sep="")
VT <- readOGR(shapefile_name, layer = "tl_2015_50_tract", verbose = FALSE)
leaflet(VT) %>%
  addTiles() %>% 
  addPolylines(color="black", weight=1)
```


### Write-Up

Upload your shiny app to the Middlebury Shiny Server (see Lecture 16) and post
the url to the app [here](https://www.nhl.com/), replacing the nhl.com link with 
the link to your app.

Comment on general ethnic demographic trends that's more substantive than just
"Vermont is really white."



